---
title: "hw6"
author: "Giang Vu"
date: "10/1/2020"
output: pdf_document
---

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(ggrepel)
library(ggplot2)
library(dplyr)
library(Hmisc)
library(readxl)
library(DBI)
library(odbc)
library(readr)
library(writexl)
library(knitr)
library(clipr)
library(car)
```

# HOMEWORK 6

## 4) 
__a)__
```{r echo=TRUE, warning=FALSE}
#read data
hw6_dt <- read.delim("/Users/giangvu/Desktop/STAT 2131 - Applied Stat Methods 1/HW/hw6/steam_text-1.txt")
#regress steam (Y) onto fat (X1) and glycerine (X2)
hw6_md <- lm(steam ~ fat + glycerine, data = hw6_dt)
#i)
hw6_sm<-summary(hw6_md) #mean model
hw6_anova <- summary(aov(hw6_md)) #variance model
hw6_sm
hw6_anova
```
Interpreting the coefficients:
As the value for variable fat increases by 1 unit, the value for steam is expected to increase by 1.728 units. As the value for variable glycerin increases by 1 unit, the value for steam is expected to decreease by -6.628 units.
However, the coefficients estimates for both variables fat and glycerin in this model are not statistically significant given alpha = 0.05


```{r echo=TRUE, warning=FALSE}
#ii) plot residual as fcn of Y^, fat, glycerine
plot(predict(hw6_md), resid(hw6_md),
     ylab  = "Residuals", xlab = "Predicted Values", main ="Residual Plot 1")
abline(0,0)

plot(hw6_dt$fat, resid(hw6_md),
     ylab  = "Residuals", xlab = "Fat", main ="Residual Plot 2")
abline(0,0)

plot(hw6_dt$glycerine, resid(hw6_md),
     ylab  = "Residuals", xlab = "Glycerin", main ="Residual Plot 3")
abline(0,0)
```
Looking at the 3 plots, I don't see any relationship between residuals and glycerin, fat, or predicted Y.

```{r echo=TRUE, warning=FALSE}
#iii) 
linearHypothesis(hw6_md,c("fat=0","glycerine=0")) #F-test
```
p-value for F test is 0.1197 > 0.05 => fail to reject H0, also I noticed that R square of this model is really small too (0.1755)
We can conclude that the two variables fat and glycerin cannot fully explain the variability of Y (steam). Their coefficients are not statistically significant.

```{r echo=TRUE, warning=FALSE}
#iv) temp against residual 
plot(hw6_dt$temp, resid(hw6_md),
     ylab  = "Residuals", xlab = "Temp", main ="Residual Plot 4")
abline(0,0)
```
There's a decreasing pattern, so we can say that residuals can be explained by temp. This is probably what makes our original model a poor model.

__b)__
```{r echo=TRUE, warning=FALSE}
#i) regress steam (Y) onto fat (X1), glycerine (X2) and temp (X3)
hw6_md2 <- lm(steam ~ fat + glycerine + temp, data = hw6_dt)
summary(hw6_md2)
linearHypothesis(hw6_md2,c("fat=0","glycerine=0"))
```
p value for F test = 0.0005569 < 0.05 => reject H0
We can say that fat and glycerine do have some impact on steam now that temp is included in the model.
However, with a new variable temp added into the model and alpha = 0.05, the coefficient estimates for fat and glycerine are sill not statistically significant (based on lm() results), but the coefficient estimate for temp is statistically significant. 

ii) Because of the inclusion of variable X3 (temp). Our original model is underfitting because we leave outt an important variable - temp, so we saw that a lot of residuals are now accounted for by temp. With temp included, we have a better model, higher R squared. 